---
title: "Dota"
author: "Nikola Shahpazov"
date: "12/3/2020"
output: html_document
---

```{r package-setup, echo = FALSE, results = 'hide', warning = FALSE, message = FALSE} 

list.of.packages <- c( "tidyverse", "caret", "Metrics", "recipes",
                       "ggplot2", "purrr", "reshape")

new.packages <- list.of.packages[(!list.of.packages %in% installed.packages() |
                                   list.of.packages %in% old.packages()[ ,"Package"])]

if (length(new.packages)) {
  install.packages(new.packages, repos = "http://cran.us.r-project.org")
}

## Load all necessary packages in RAM (working environment)
lapply(list.of.packages, require, character.only = TRUE)

## Setup working directory to the path of the source file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
rm(list.of.packages,new.packages)
```

```{r constants}
WINDOW_COLS <- c("gold_difference", "xp_difference", "lh_difference")
```

```{r load-inner-modules}
source("helpers.R")
source("markov-chains-helpers.R")
```

### Loading data
```{r load-data}
matches_df <- read.csv("./data/match.csv",
                       colClasses = c(radiant_win = "logical"))
time_df <- read.csv('./data/player_time.csv')
```

### Preprocessing

```{r prepare-data echo = FALSE}
diffs_by_time_df <- time_df %>%
  mutate(gold_difference = diff_means(., "gold_t_[0-4]$", "gold_t_1[2-3]")) %>%
  mutate(xp_difference = diff_means(., "xp_t_[0-4]$", "xp_t_1[2-3]")) %>%
  mutate(lh_difference = diff_means(., "lh_t_[0-4]$", "lh_t_1[2-3]")) %>%
  mutate(minute = times / 60) %>%
  select(match_id, minute, gold_difference, xp_difference, lh_difference) %>% 
  pivot_wider(id_cols = "match_id",
              names_from = "minute",
              values_from = WINDOW_COLS) %>%
  select(-gold_difference_0, -xp_difference_0, -lh_difference_0) %>% 
  left_join(select(matches_df, match_id, radiant_win), by = "match_id") %>%
  mutate(radiant_win = as.factor(radiant_win))
```

### Train set and Test set split

```{r test-train-split echo = FALSE}
train_indices <- createDataPartition(diffs_by_time_df$radiant_win,
                                     p = .8,
                                     list = FALSE,
                                     times = 1)

train_set <- diffs_by_time_df[train_indices, ]
test_set <- diffs_by_time_df[-train_indices, ]

```

### Train various logistic regression models

```{r train-time-windows-lr echo = FALSE}
tc <- trainControl(method = "cv", number = 5)

train_set_std <- standardize_time_df(train_set)
test_set_std <- standardize_time_df(test_set)

# slr_all <- train_lr(data = train_set_std, cols = WINDOW_COLS, control = tc)
slr_gld <- train_lr(data = train_set_std, cols = WINDOW_COLS[1], control = tc)
slr_glh <- train_lr(data = train_set_std, cols = WINDOW_COLS[-2], control = tc)

lr_models <- list(
  # "slr_all" = slr_all,
  "slr_gld" = slr_gld,
  "slr_glh" = slr_glh
)
```

### Explore Cross Validation Test Accuracy

```{r plot-cv-acuracy}
lr_models %>%
  map(~imap(., ~.$results$Accuracy)) %>%
  imap(~data.frame(accuracy = unlist(.), time = 6:101, model = .y)) %>%
  bind_rows() %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Training set CV Accuracy")
```

We see that SLR with gold and last hits (gold is correlated with xp) behaves the best till the 
80th minute and after that it starts oscillating a lot, probably because 
there aren't a lot of matches then, or the matches are very even.

```{r prediction-accuracies echo = FALSE}
test_performance <- lr_models %>%
  imap(
    ~data.frame(
      accuracy = get_test_accuracies(., test_set_std),
      time = 5:95,
      model = .y
    )
  ) %>%
  bind_rows()

test_performance %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Test Set Accuracy")
```
We see that the test accuracy behaves similarly to the test set cv accuracy from above.


### Markov Chains Model

```{r markov-chain-model}
train_set %>%
  filter(as.logical(radiant_win)) %>%
  as_binned_matrix() %>%
  construct_transition_matrix() %>%
  matrix_to_df() %>%
  ggplot(mapping = aes(col, row, fill = value)) + 
    geom_tile() +
    ggtitle("Transition changes for the Winning team in the train set") +
    xlab("States") +
    ylab("States")
```