---
title: "Dota"
author: "Nikola Shahpazov"
date: "12/3/2020"
output: html_document
---
```{r}
install.packages("tidyverse")
install.packages("caret")
install.packages("gridExtra")
install.packages("reshape2")
```

```{r package-setup, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library("tidyverse")
library("caret")
library("gridExtra")
library("reshape2")
```

```{r constants}
WINDOW_COLS <- c("gold_difference", "xp_difference", "lh_difference")
NBIN <- 12
```

```{r load-inner-modules}
source("helpers.R")
source("markov-chains-helpers.R")
```

### Loading data
```{r load-data}
matches_df <- read.csv(
  "./data/match.csv",
  colClasses = c(radiant_win = "logical")
)

time_df <- read.csv('./data/player_time.csv')
```

### Preprocessing

```{r prepare-data}
diffs_by_time_df <- time_df %>%
  mutate(minute = times / 60) %>%
  filter(minute <= 100) %>%
  mutate(gold_difference = diff_means(., "gold_t_[0-4]$", "gold_t_1[2-3]")) %>%
  mutate(xp_difference = diff_means(., "xp_t_[0-4]$", "xp_t_1[2-3]")) %>%
  mutate(lh_difference = diff_means(., "lh_t_[0-4]$", "lh_t_1[2-3]")) %>%
  select(match_id, minute, gold_difference, xp_difference, lh_difference) %>%
  pivot_wider(id_cols = "match_id", names_from = "minute",
              values_from = WINDOW_COLS) %>%
  select(-gold_difference_0, -xp_difference_0, -lh_difference_0) %>%
  left_join(select(matches_df, match_id, radiant_win),
            by = "match_id") %>%
  mutate(radiant_win = as.factor(radiant_win))
```

### Train set and Test set split

```{r test-train-split}
train_indices <- createDataPartition(
  diffs_by_time_df$radiant_win,
  p = .8,
  list = FALSE,
  times = 1
)

train_set <- diffs_by_time_df[train_indices, ]
test_set <- diffs_by_time_df[-train_indices, ]
```

### Train various logistic regression models

```{r train-time-windows-lr, include=FALSE}
tc <- trainControl(method = "cv", number = 3)

# Стандартизираме данните
train_set_std <- standardize_time_df(train_set) %>%
  filter(is.na(gold_difference_50))
test_set_std <- standardize_time_df(test_set) %>%
  filter(is.na(gold_difference_50))

slr_gld <- train_lr(data = train_set_std, cols = WINDOW_COLS[1], control = tc)
slr_glh <- train_lr(data = train_set_std, cols = WINDOW_COLS[-2], control = tc, interval = 0:50)
slr_ge <- train_lr(data = train_set_std, cols = WINDOW_COLS[-3], control = tc)
slr_all <- train_lr(data = train_set_std, cols = WINDOW_COLS, control = tc)

dt_all <- train_dt(data = train_set_std, cols = WINDOW_COLS, control = tc)

lr_models <- list(
  # "Злато" = slr_gld,
  # "Злато и Опит" = slr_ge,
  "Злато и Удари" = slr_glh
  # "Злато, Опит и Удари" = slr_all
)

```

### Explore Cross Validation Test Accuracy

```{r plot-cv-acuracy}
lr_accuracies <- lr_models %>%
  map(~imap(., ~.$results$Accuracy)) %>%
  imap(~data.frame(accuracy = unlist(.), time = 6:39, model = .y)) %>%
  bind_rows() 

g <- lr_accuracies %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Трениране на различни модели") +
  xlab("Време") +
  ylab("Точност") +
  theme(legend.position = "bottom")

g

ggsave(
  "images/real-time-predictors.png",
  g,
  width = 10,
  height = 6
)
```

We see that SLR with gold and last hits (gold is correlated with xp) behaves the best till the 
80th minute and after that it starts oscillating a lot, probably because 
there aren't a lot of matches then, or the matches are very even.

```{r prediction-accuracies}
# test_performance <- lr_models %>%
#   imap(
#     ~data.frame(
#       accuracy = get_test_accuracies(., test_set_std, interval = 1:60),
#       time = 1:60,
#       model = .y
#     )
#   ) %>%
#   bind_rows()

library(Metrics)

get_test_accuracies(lr_accuracies)

test_performance %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Test Set Accuracy")
```
We see that the test accuracy behaves similarly to the test set cv accuracy from above.

### Markov Chains Model

```{r markov-chain-model, message=FALSE, warning=FALSE}
# create winners transition matrix
NBIN <- 60
MINUTES <- 1:40

train_set2 <- train_set %>%
  select(match_id, radiant_win, starts_with("gold")) %>%
  # filter(!is.na(gold_difference_40)) %>%
  filter(is.na(gold_difference_42)) %>%
  .[, 1:42]

test_set2 <- test_set %>%
  select(match_id, radiant_win, starts_with("gold")) %>%
  # filter(!is.na(gold_difference_40)) %>%
  filter(is.na(gold_difference_42)) %>%
  .[, 1:42]

winners_tm <- train_set2 %>%
  filter(as.logical(radiant_win)) %>%
  as_binned_matrix(nbin = NBIN, minutes = MINUTES) %>%
  construct_transition_matrix(nbin = NBIN)

# TODO: think of some way to use turnouts (strong jumps in gold)
# TODO: explore, what are the number of turnouts for winning teams?
# TODO: add kills per min

# create losers transition matrix
losers_tm <- train_set2 %>%
  filter(!as.logical(radiant_win)) %>%
  as_binned_matrix(nbin = NBIN, minutes = MINUTES) %>%
  construct_transition_matrix(nbin = NBIN)

test_set_bin_matrix <- as_binned_matrix(test_set2, nbin = NBIN, minutes = MINUTES)
# tsbm_pca <- pca_test_set %>% as_binned_matrix(nbin = NBIN, feature = "pc")

# pca_mc_accuracies <- data.frame(
#   accuracy = unlist(imap(1:80, ~acc(test_set, tsbm_pca, winners_pca_tm, losers_pca_tm, .x))),
#   time = 1:80,
#   model = "MC PCA"
# )

# test_set_bin_matrix[, 40]

# get_mc_accuracy(test_set2, test_set_bin_matrix, winners_tm, losers_tm, 35)

mc_accuracies <- data.frame(
  accuracy = unlist(
    imap(6:39, ~get_mc_accuracy(test_set2, test_set_bin_matrix, winners_tm, losers_tm, .x))
  ),
  time = 6:39,
  model = "Марковска Верига"
)

g14 <- data.frame(
  model = "Логистична Регресия със Злато и Удари",
  accuracy = unlist(map(slr_glh, ~.x$results$Accuracy))[6:40],
  time = 6:40
) %>%
  bind_rows(mc_accuracies) %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Точност по време на МВ и ЛР") +
  xlab("Време") +
  ylab("Точност") +
  theme(legend.position = "bottom")

g14
```


```{r markov-chain-model, message=FALSE, warning=FALSE}
# plot winners transition matrix
g1 <- winners_tm %>%
  matrix_to_df() %>%
  ggplot(mapping = aes(row, col, fill = value)) +
    geom_tile() +
    # geom_text(aes(label = round(value, 1))) +
    ggtitle("Емпирични Преходни Вероятности за Победителя") +
    xlab("До Клетка със Злато (j)") +
    ylab("От Клетка със Злато (i)")

# plot losers transition matrix
g2 <- losers_tm %>%
  matrix_to_df() %>%
  ggplot(mapping = aes(row, col, fill = value)) + 
    geom_tile() +
    # geom_text(aes(label = round(value, 1))) +
    ggtitle("Емпирични Преходни Вероятности за Губещия") +
    xlab("До Клетка със Злато (j)") +
    ylab("От Клетка със Злато (i)")

g3 <- grid.arrange(g1, g2, ncol = 2)
g1
g2

ggsave(
  "images/markov-chain-transition-probabilities-60.png",
  g3,
  width = 10,
  height = 6
)

```


```{r markov-chains-model-predictions, echo=FALSE}
test_set_bin_matrix <- as_binned_matrix(test_set2, nbin = NBIN, minutes = MINUTES)
# tsbm_pca <- pca_test_set %>% as_binned_matrix(nbin = NBIN, feature = "pc")

# pca_mc_accuracies <- data.frame(
#   accuracy = unlist(imap(1:80, ~acc(test_set, tsbm_pca, winners_pca_tm, losers_pca_tm, .x))),
#   time = 1:80,
#   model = "MC PCA"
# )

# test_set_bin_matrix[, 40]

# get_mc_accuracy(test_set2, test_set_bin_matrix, winners_tm, losers_tm, 35)

mc_accuracies <- data.frame(
  accuracy = unlist(
    imap(6:39, ~get_mc_accuracy(test_set2, test_set_bin_matrix, winners_tm, losers_tm, .x))
  ),
  time = 6:39,
  model = "Марковска Верига"
)

# benchmark_accuracies <- test_set2 %>%
#   mutate_at(vars(starts_with("gold")), ~as.numeric((. > 0) == radiant_win)) %>%
#   summarise_at(vars(starts_with("gold")), mean, na.rm = TRUE) %>%
#   pivot_longer(cols = gold_difference_6:gold_difference_39, values_to = "accuracy") %>%
#   mutate(time = 6:39, model = "Benchmark")

# TODO: train logistic regression only on 40 minutes matches and plot again to see the difference

# bind_rows(
#   # test_performance[1:35, ],
#   mc_accuracies,
#   benchmark_accuracies
# ) %>%
# ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
# geom_line() +
# ggtitle("Test Set Accuracy")
```


```{r markov-chains-model-predictions, echo=FALSE}
g_logistic_and_mc <- lr_accuracies %>%
  filter(model == "Злато и Удари") %>%
  mutate(model = "Логистична Регресия със Злато и Удари") %>%
  rbind("Марковска Верига" = mc_accuracies) %>%
  filter(time <= 80) %>%
  ggplot(mapping = aes(x = time, y = accuracy, color = model)) +
  geom_line() +
  ggtitle("Точност по време на МВ и ЛР") +
  xlab("Време") +
  ylab("Точност") +
  theme(legend.position = "bottom")


g_logistic_and_mc

ggsave(
  "images/lr-and-mc.png",
  g_logistic_and_mc,
  width = 10,
  height = 6
)
```


### Notes

* Tried reflecting losers and winners but it brought little to the table

* We tried segmenting the markov chain models into different categories. When predicting
for a minute we use a model for which is trained on games longer than that minute
This didn't improve anything, probably because the gold transition changing is more
important in early stages of the game.

* Tried different quantile bins for the gold states and seems that 8 gives better results than 10

* Tried PCA on a transformation of gold and last hits but the Markov Chain seems to be performing slightly worse than the original

### Markov Chain model with a PCA-generated state space

```{r markov-pca, include=FALSE}
pca_data <- diffs_by_time_df %>%
  standardize_time_df() %>%
  pca_transform(cols = WINDOW_COLS[-2])

pca_data$match_id <- diffs_by_time_df$match_id
pca_data$radiant_win <- diffs_by_time_df$radiant_win

pca_train_set <- pca_data[train_indices, ]
pca_test_set <- pca_data[-train_indices, ]

winners_pca_tm <- pca_train_set %>%
  filter(as.logical(radiant_win)) %>%
  as_binned_matrix(nbin = NBIN, feature = "pc") %>%
  construct_transition_matrix(<- = NBIN)

# create losers transition matrix for pca data
losers_pca_tm <- pca_train_set %>%
  filter(!as.logical(radiant_win)) %>%
  as_binned_matrix(nbin = NBIN, feature = "pc") %>%
  construct_transition_matrix(nbin = NBIN)

winners_pca_tm %>%
  matrix_to_df() %>%
  ggplot(mapping = aes(row, col, fill = value)) + 
    geom_tile() +
    geom_text(aes(label = round(value, 1))) +
    ggtitle("Transition changes for the winning team in the train set") +
    xlab("PCA Bin States") +
    ylab("PCA Bin States")

losers_pca_tm %>%
  matrix_to_df() %>%
  ggplot(mapping = aes(row, col, fill = value)) + 
    geom_tile() +
    geom_text(aes(label = round(value, 1))) +
    ggtitle("Transition changes for the losing team in the train set") +
    xlab("PCA Bin States") +
    ylab("PCA Bin States")
```